Prospects of detecting gamma-ray emission from galaxy clusters: cosmic rays and dark matter annihilations

### 2011-03-30 paper.tex
updated figures+theory

%NEW
\newcommand{\rmn}{\mathrm}
\newcommand{\bx}{{\bf x}}
\newcommand{\fg}{{F_\gamma}}
\newcommand{\sg}{{S_\gamma}}
\newcommand{\psf}{\theta_\rmn{res}}
\newcommand{\ph}{\rmn{ph}}
\newcommand{\eph}{E_\ph}
\newcommand{\nph}{N_\ph}
\newcommand{\vir}{\rmn{vir}}
\newcommand{\gal}{\rmn{gal}}
\newcommand{\sd}{\rmn{SD}}
\newcommand{\clu}{\rmn{cl}}
\newcommand{\sfe}{\rmn{sfe}}
\newcommand{\sub}{\rmn{sub}}
\newcommand{\msun}{M_\odot}
\newcommand{\ee}{E_\rmn{e}}
\newcommand{\stars}{\rmn{stars}}
\newcommand{\dust}{\rmn{dust}}
\newcommand{\lx}{L_\rmn{X}}
\newcommand{\s}{\rmn{s}}

%OLD
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\gev}{\rm{GeV}}
\newcommand{\tev}{\rm{TeV}}
\newcommand{\ev}{\rm{eV}}
\newcommand{\dd}{\rmn{d}}
\newcommand{\mx}{\ensuremath{m_{\chi}}}
\newcommand{\ga}{\gamma}
\newcommand{\ngamma}{\ensuremath{N_{\gamma}}}
\newcommand{\ngammaj}{\ensuremath{N_{\gamma,j}}}
\newcommand{\sigmaannv}{\ensuremath{\langle\sigma v\rangle}}
\newcommand{\sigv}{\ensuremath{\sigma_v}}
\newcommand{\ic}{\rmn{IC}}
\newcommand{\egamma}{\ensuremath{E_{\gamma}}}
\newcommand{\dic}{\rmn{dIC}}
\newcommand{\fsr}{\rmn{FSR}}
\newcommand{\CR}{\rmn{CR}}
\newcommand{\rhos}{\ensuremath{\rho_s}}
\newcommand{\rs}{\ensuremath{r_s}}
\newcommand{\rvir}{r_{200}}
\newcommand{\mvir}{M_{200}}
\newcommand{\rhoc}{\ensuremath{\rho_c}}
\newcommand{\B}{\rmn{B}}
\newcommand{\e}{\rmn{e}}
\newcommand{\eg}{E_\gamma}
\newcommand{\ir}{\rmn{ir}}
\newcommand{\Km}{\rmn{K}`}
\newcommand{\Imm}{\rmn{I}`}
\newcommand{\Jm}{\rmn{J}^*}
\newcommand{\Jmm}{\rmn{J}`}

@@

\rho(r) = \frac{\rho_\s}{\left(r/r_\s+\right)^\gamma

@@

and the characteristic density by $\rhos=(\Delta\rhoc/3)\,
c^3/(\log(1+c)-(c/(1+c)))$, where the halo mass dependent
concentration parameter $c$ is derived from a power-law fit to
cosmological simulations with $\mvir \gtrsim 10^{10} \msun$
\cite{2008MNRAS.391.1940M},

@@

Recent detailed galaxy cluster observations find a core like density
profile with an inner slope of $\beta=0.6$, where $\beta\gtrsim 1$ can
be ruled out to a $>95$\% confidence \cite{2011ApJ...728L..39N}. This
finding is hard to reconcile with simulations that typically prefere a
steeper central profile. The different density profiles have a large
impact on the luminosity from annihilating DM, although the details of
the density profile can be neglected compared to the dominating boost
from substructures. In Fig.~\ref{fig:radial_lum} we compare the radial
dependence of the accumulative luminosity from different smooth
cluster density profiles to the substructure boosted luminosity at
different mass scales. We recalculate the density contrast for the
core like density profile $\rhos$ and rescale the concentration
parameter in Eq.~(\ref{eq:cfit}) with $300/160$
\cite{2011ApJ...728L..39N} to account for the more centralized scale
radius in a core like cluster. The emission coming from the core
compared to a cuspy like density profile is about 30\% larger within
$\rvir$. This difference is build up within 10\% of $\rvir$
(i.e. close to $r_\rmn{s}$). The emission from an Einasto density
profile is about 50\% larger than the cuspy NFW profile in the
periphery of the cluster, where the difference is mainly build up at a
few percent of $\rvir$. The substructure boosted luminosities is
neglible in the center of halos, but integrated out to $\rvir$ the
boost is about; 10 for dwarf galaxies, $10^2$ for galaxies, and $10^3$
for galaxy clusters.
\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/dens.prof.eps}
\caption{\it Radial dependence of various luminosity
  contributions. The solid lines show the accumulative smooth
  luminosity from a cluster with the mass $\mvir=10^{15}\,\msun$ for
  three different density profiles; an Einasto profile with
  $\alpha=0.17$ (light blue), a cuspy NFW profile with $\beta=1.0$
  (dark blue), and a core NFW profile with $\beta=0.6$ (purple). The
  dashed lines show the accumulative luminosity from substructures for
  three different mass scales; an $\mvir=10^{15}\,\msun$ galaxy
  cluster (red), an $\mvir=10^{12}\,\msun$ galaxy (orange), and an
  $\mvir=10^{8}\,\msun$ dwarf galaxy (green). All luminosities have
  been normalized with the luminosity within $\rvir$ from a cuspy NFW
  profile. We have assumed the standard value for the limiting
  substructure mass of $10^{-6}\,\msun$. Note the large expected
  boost from substructures in clusters ($\sim1000$), and the relative
  small boost in dwarf galaxies ($\sim20$).}
 \label{fig:radial_lum}
\end{figure}

@@

\subsubsection{Cosmic-ray induced gamma-ray emission}
\label{sect:CRs}
The formation process of a galaxy cluster is a very energetic
processes that induces both turbulence as well as frequently occurring
merging and accretion shocks. These processes are thought to
accelerate large quantities of relativistic non-thermal electrons and
protons to high energies. On smaller scales, such as galaxies, there
is convincing evidence of such non-thermal populations. Especially, in
the MW, the cosmic rays are observed directly as well as indirectly
through radio, X-ray, and gamma-ray emission. On larger scales of the
order of few 100 kpc up to Mpc, there are a vast number of
observations of radio emission coming from radio mini halos in the
center of cooling flow clusters, radio relics in the periphery of
clusters \cite{2004rcfg.proc..335K} as well smooth giant radio halos
that have been observed in more than 50 clusters
\cite{2003ASPC..301..143F,2008SSRv..134...93F}. Most of this observed
radio emission is expected to emerge from synchrotron emitting
cosmic-ray electrons whose precise origin in especially relics and
giant radio halos is still not settled. However, due to the smoothness
of the extended emission and the observed power-law slope it seems
quite natural that they are of hadronic origin. Especially since the
long cooling time of cosmic ray protons (CRs) allows for both an
extended and large population of CRs to build up over the cluster
history \cite{1997ApJ...487..529B}.

@@

\subsubsection{Inverse Compton}
\label{sect:IC}
The standard IC source function is given by
\cite{1979rpa..book.....R}:
\begin{eqnarray}
  q_{\rmn IC}(\eg, r) &=&  \frac{\dd^3 N_\gamma}{\dd V\,\dd t\,\dd \eg} = 
 \frac{3}{4}c\,\sigma_\rmn{T}
\int\dd \eph \frac{n_\rmn{ph}(\eph)}{\eph}\nonumber\\
&\times& \int \dd \ee\,\frac{\dd n_\e}{\dd \ee}(\ee,r)\,
 \frac{\left(m_\e c^2\right)^2}{\ee^2}G(\Gamma_\e,q)\,,\nonumber\\
  \label{eq:ICemiss}
\end{eqnarray}
where $\ee$ is the energy of the upscattering electrons and $\eph$ is
the energy of the background photon field. We represent the Thomson
cross-section with $\sigma_\rmn{T}$ and the full differential
Klein-Nishina (KN) cross-section is captured by
\cite{1970RvMP...42..237B}:
\begin{equation}
\label{eq:KN_spec}
G(\Gamma_\e,q) = 2q\ln{q}+(1+2q)(1-q)+ 
\frac{1}{2}\frac{\left(\Gamma_\e q\right)^2\left(1-q\right)}
     {1+\Gamma_\e q}\,,
\end{equation}
where
\begin{equation}
\Gamma_\e=\frac{4\eph \ee}{\left(m_\e c^2\right)^2}\,,\qquad \rmn{and} \qquad  
q=\frac{\eg}{\Gamma_\e\left(\ee-\eg\right)}\,.
\end{equation}
The full KN cross-section accounts for the less efficient energy
transfer between the photon and electron and results in a steepening
of the IC gamma-ray spectrum. In the low energy Thomson regime the IC
spectrum is given by $E_\gamma^{-(\alpha_\e-1)/2}$, however when
$\Gamma_\e \gg 1$ the IC spectrum steepens due to the KN suppression
to $\eg^{-\alpha_\e}\log(\eg)$ .

@@

\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/SB.photon.eps}
\caption{\it Spatial dependence of light from stars and dust. We show
  2D surface brightness profiles in the $r$-band ($\sim 1\,\ev$)
  obtained from stacked clusters in the Sloan Digital Sky Survey
  (SDSS) at the redshift $z \sim 0.25$ \cite{2005MNRAS.358..949Z}. The
  brightness profiles have been weighted with the $\rvir$ normalized
  area inside radius $r$, and trace the radial dependence of the
  luminosity from stars and galaxies. The crosses show the total
  observed star light including the diffuse intracluster (ICL), the
  galaxies, and the brightest cluster galaxy (BCG) in the center of
  the cluster. The diamonds denote the observed star light from the
  ICL and the BCG. The solid line show the fit to the data of the
  total light including the ICL, the BCG, and the galaxies, while the
  smooth component is represented by the dashed line that is fitted to
  only the ICL and the BCG.}
 \label{fig:SD_spatial}
\end{figure}

@@

The normalization constant for the star light [$N_\stars(\lx)$] is derived
in the following three steps: (1) We use the average apparent magnitude
in the $(r+i)$-band $m_{(r+i),0.25}\approx 15.5$
\cite{2005MNRAS.358..949Z} of a cluster with the average mass
$\mvir=4.0\times10^{13}\msun$ at a redshift $z=0.25$ to derive the
luminosity from the stars:
\begin{equation}
L_\stars=10^{\frac{m_\odot-m_{(r+i),0.25}}{2.5}}
\left(\frac{d_\clu}{d_\odot}\right)^2 L_\odot
\approx 7.1\times10^{44}\,\rmn{erg}\,\rmn{s}^{-1}\,,
\label{eq:L_star}
\end{equation}
where the apparent magnitude of the sun $m_\odot\approx -27.1$, the
distance to the sun $d_\odot=4.85\times10^{-6}\,\rmn{pc}$, the
luminosity of the sun $L_\odot=3.85\times10^{34}\,\rmn{erg\,s}^{-1}$,
and the distance to the cluster $d_\clu=1.26\times10^9\,\rmn{pc}$. (2)
We use the IR to X-ray scaling relation in Eq.~(\ref{eq:SD-X}) to find
a more general form of the luminosity from stars in
Eq.~(\ref{eq:L_star}) that is valid for an arbitrary cluster mass. The
cluster X-ray luminosities are found in the extended HIFLUGCS
catalogue \cite{2002ApJ...567..716R} to which we have added the Virgo
cluster. (3) The normalization constant for the stars is
determined from the above steps together with Eq.~(\ref{eq:E_SD})
where we integrate over the cluster volume and spectral distribution
of the stars. It is given by
\begin{equation}
 N_\stars(\lx) = 1.2\times10^{-9}\,
\left(\frac{\lx}{10^{44}\rmn{erg\,s}^{-1}}\right)^{0.53}\,.
\label{eq:N_stars}
\end{equation}
Note that since the luminosities include the contribution from
galaxies, we have integrated the spatial distribution including
galaxies to derive the normalization. Once the normalization is fixed,
the SD energy densities are derived from the spatial distribution
where galaxies have been excluded.

The normalization constant for the dust [$N_\dust(\lx)$] is derived in the
following four steps: (1) We use a cluster with the the average mass
$\mvir=4.0\times10^{13}\msun$ and estimate the X-ray luminosity from
the X-ray to mass scaling relations derived for the Representative
XMM-Newton Cluster Structure Survey (REXCESS) in
\cite{2009A&A...498..361P}:
\begin{equation}
\lx=1.18\times 10^{44}\,\rmn{erg\,s}^{-1}
\left(\frac{M_{200}}{10^{14}\msun}\right)^{1.81}\,,
\end{equation}
where we have used that $M_\rmn{200} \approx 1.58\,M_\rmn{500}$ for a
small cluster \footnote{We derive the factor 1.58 using
  $M_\rmn{200}/M_\rmn{500}$ for the clusters in the HIFLUGCS catalogue
  with a mass similar to $\mvir=4.0\times10^{13}\msun$ that we use to
  derive the luminosity from dust.}. (2) From the IR to X-ray scaling
relation in Eq.~(\ref{eq:SD-X}) we can now estimate calculate the
luminosity from the dust for a $\mvir=4.0\times10^{13}\msun$ cluster;
$L_\dust=4.7\times10^{44}\,\rmn{erg\,s}^{-1}$. (3) We generalize
Eq.~(\ref{eq:L_star}) to an arbitrary cluster mass using
Eq.~(\ref{eq:SD-X}) together with the cluster X-ray luminosities in
the extended HIFLUGCS catalogue. (4) The normalization constant for
the dust can now be determined from from the above steps together with
Eq.~(\ref{eq:E_SD}) where we integrate over the cluster volume and
spectral distribution of the dust. It is given by
\begin{equation}
 N_\dust(\lx) = 1.4\times10^{-9}\,
\left(\frac{\lx}{10^{44}\rmn{erg\,s}^{-1}}\right)^{0.53}\,.
\label{eq:N_stars}
\end{equation}
Note that since the luminosities include the contribution from
galaxies, we have integrated the spatial distribution including
galaxies to derive the normalization. Once the normalization is fixed,
the SD energy densities are derived from the spatial distribution
where galaxies have been excluded.

@@

\section{Spectral gamma-ray distribution}

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
 \includegraphics[width=0.49\columnwidth]{figures/flux.BMcompI.v9.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompJ.v9.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompK.v9.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompJs.v9.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\caption{\it Comparing the differential flux from different models: we
  show the continuum emission from DM benchmark (BM) models
  (blue), electrons and positrons from DM BM models that inverse
  Compton upscattered both CMB photons (orange) as well as dust and
  star photons (green), and CR induced gamma-ray emission (red
  solid). The various linestyles are associated with individual DM BM
  models; $\Imm$ (dotted), $\Km$ (dashed), and $\Jm$
  (dash-dotted). The emission is calculated for the Fornax cluster
  using a point spread function of $0.1\deg$ and a boost from
  substructures of ...  We find bright prospects for detecting the
  continuum emission that is dominating the total emission in the GeV
  energy range. Also note that, above 100 MeV the total inverse
  Compton emission is at least a factor 10 smaller than the emission
  expected from both the continuum emission and the emission coming
  from CRs.}
 \label{fig7}
\end{minipage}
\end{figure*}

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
\includegraphics[width=0.49\columnwidth]{figures/flux.IRcomp.v9.0.1deg.1.6T.SubMass.elmu.SF300.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.IRcomp.BMv9.0.1deg.SubMass.noMW.woGal.eps}
\caption{\it Comparing the differential flux from different inverse
  Compton upscattered radiation fields in the Fornax cluster. We show
  the inverse Compton emission induced by leptophilic DM in the left
  panel and by the $\Km$ benchmark model in the right panel. The
  contribution from each individual radiation field from top line to
  bottom line; CMB (green dashed-double-dotted), dust (blue dashed),
  stars (purple dashed-dotted). The sum of the three components are
  shown with the orange solid line. The black dotted line represents
  the upscattered CMB photons without any cooling from stars and
  dust. The red solid line show the CR induced gamma-ray flux. The
  black arrow show the differential upper limit from Fermi ADD
  REF. All fluxes are calculated within $\rvir$ with a point spread
  function of $0.1\deg$. The boost from Sommerfeld is ..., and from
  substructures is .}
 \label{fig6}
\end{minipage}
\end{figure*}

\begin{figure}
 \includegraphics[width=0.99\columnwidth]{figures/flux.int.v9.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\caption{\it Comparing the energy integrated flux from different
  models: we show the emission from leptophilic (LP) models with green
  lines, $\Km$ benchmark model (BM) with blue lines, and the CR
  induced emission with a red line. The dotted and dashed lines show
  the inverse Compton (IC) upscattered CMB photons and star and dust
  photons, respectively. The green solid line show the final state
  radiation from the LP model, and the solid blue line show the
  continuum emission from the DM. The black arrow show the integrated
  flux upper limit set by Fermi.  The emission is calculated for the
  Fornax cluster using a point spread function of $0.1\deg$ and a
  boost from substructures of ...  We find that LP models are
  dominating the total flux in the entire energy range, where the IC
  upscattered CMB photons overproduce the upper limit by a factor 3.}
 \label{fig8}
\end{figure}

>>>>

%NEW
\newcommand{\rmn}{\mathrm}
\newcommand{\bx}{{\bf x}}
\newcommand{\fg}{{F_\gamma}}
\newcommand{\sg}{{S_\gamma}}
\newcommand{\psf}{\theta_\rmn{res}}
\newcommand{\ph}{\rmn{ph}}
\newcommand{\eph}{E_\ph}
\newcommand{\nph}{N_\ph}
\newcommand{\vir}{\rmn{vir}}
\newcommand{\gal}{\rmn{gal}}
\newcommand{\sd}{\rmn{SD}}
\newcommand{\clu}{\rmn{cl}}
\newcommand{\sfe}{\rmn{sfe}}
\newcommand{\sub}{\rmn{sub}}
\newcommand{\msun}{M_\odot}
\newcommand{\ee}{E_\rmn{e}}
\newcommand{\stars}{\rmn{stars}}
\newcommand{\dust}{\rmn{dust}}
\newcommand{\lx}{L_\rmn{X}}
\newcommand{\s}{\rmn{s}}
\newcommand{\Kp}{\rmn{K}`}
\newcommand{\Ip}{\rmn{I}`}
\newcommand{\Js}{\rmn{J}^*}
\newcommand{\Jp}{\rmn{J}`}

%OLD
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\gev}{\rm{GeV}}
\newcommand{\tev}{\rm{TeV}}
\newcommand{\ev}{\rm{eV}}
\newcommand{\dd}{\rmn{d}}
\newcommand{\mx}{\ensuremath{m_{\chi}}}
\newcommand{\ga}{\gamma}
\newcommand{\ngamma}{\ensuremath{N_{\gamma}}}
\newcommand{\ngammaj}{\ensuremath{N_{\gamma,j}}}
\newcommand{\sigmaannv}{\ensuremath{\langle\sigma v\rangle}}
\newcommand{\sigv}{\ensuremath{\sigma_v}}
\newcommand{\ic}{\rmn{IC}}
\newcommand{\egamma}{\ensuremath{E_{\gamma}}}
\newcommand{\dic}{\rmn{dIC}}
\newcommand{\fsr}{\rmn{FSR}}
\newcommand{\CR}{\rmn{CR}}
\newcommand{\rhos}{\ensuremath{\rho_s}}
\newcommand{\rs}{\ensuremath{r_s}}
\newcommand{\rvir}{r_{200}}
\newcommand{\mvir}{M_{200}}
\newcommand{\rhoc}{\ensuremath{\rho_c}}
\newcommand{\B}{\rmn{B}}
\newcommand{\e}{\rmn{e}}
\newcommand{\eg}{E_\gamma}
\newcommand{\ir}{\rmn{ir}}

@@

\rho(r) = \frac{\rhos}{\left(r/r_\s+\right)^\gamma

@@

and the characteristic density by 
\rhos(c)=\frac{\Delta\,\rhoc}{3}\,\frac{c^3}
{\log\left(1+c\right)-c/(1+c)}\,,
\label{eq:rho_s}
 where the halo mass dependent concentration parameter $c$ is derived
 from a power-law fit to cosmological simulations with $\mvir \gtrsim
 10^{10} \msun$ \cite{2008MNRAS.391.1940M},

@@

Recent detailed galaxy cluster observations find a core like density
profile with an inner slope of $\beta=0.6$, where $\beta\gtrsim 1$ can
be ruled out to a $>95$\% confidence \cite{2011ApJ...728L..39N}. This
finding is hard to reconcile with simulations that typically prefere a
steeper central profile. The different density profiles have a large
impact on the luminosity from annihilating DM, although the details of
the density profile can be neglected compared to the dominating boost
from substructures. In Fig.~\ref{fig:radial_lum} we compare the radial
dependence of the accumulative luminosity from different smooth
cluster density profiles to the substructure boosted luminosity at
different mass scales. We recalculate the density contrast for the
core like density profile $\rhos$ and rescale the concentration
parameter in Eq.~(\ref{eq:cfit}) with $300/160$
\cite{2011ApJ...728L..39N} to account for the more centralized scale
radius in a core like cluster. The emission coming from the core
compared to a cuspy like density profile is about 30\% larger within
$\rvir$. This difference is build up within 10\% of $\rvir$
(i.e. close to $r_\s$). The emission from an Einasto density
profile is about 50\% larger than the cuspy NFW profile in the
periphery of the cluster, where the difference is mainly build up at a
few percent of $\rvir$. The substructure boosted luminosities is
neglible in the center of halos, but integrated out to $\rvir$ the
boost is about; 10 for dwarf galaxies, $10^2$ for galaxies, and $10^3$
for galaxy clusters.
\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/dens.prof.eps}
\caption{\it Radial dependence of various luminosity
  contributions. The solid lines show the accumulative smooth
  luminosity from a cluster with the mass $\mvir=10^{15}\,\msun$ for
  three different density profiles; an Einasto profile with
  $\alpha=0.17$ (light blue), a cuspy NFW profile with $\beta=1.0$
  (dark blue), and a core NFW profile with $\beta=0.6$ (purple). The
  dashed lines show the accumulative luminosity from substructures for
  three different mass scales; an $\mvir=10^{15}\,\msun$ galaxy
  cluster (red), an $\mvir=10^{12}\,\msun$ galaxy (orange), and an
  $\mvir=10^{8}\,\msun$ dwarf galaxy (green). All luminosities have
  been normalized with the luminosity within $\rvir$ from a cuspy NFW
  profile. We have assumed the standard value for the limiting
  substructure mass of $10^{-6}\,\msun$. Note the large expected
  boost from substructures in clusters ($\sim1000$), and the relative
  small boost in dwarf galaxies ($\sim20$).}
 \label{fig:radial_lum}
\end{figure}

@@

\subsubsection{Cosmic-ray induced gamma-ray emission}
\label{sect:CRs}
The formation process of a galaxy cluster is a very energetic
processes that induces both turbulence as well as frequently occurring
merging and accretion shocks. These processes are thought to
accelerate large quantities of relativistic non-thermal electrons and
protons to high energies. On smaller scales, such as galaxies, there
is convincing evidence of such non-thermal populations. Especially, in
the MW, the cosmic rays are observed directly as well as indirectly
through radio, X-ray, and gamma-ray emission. On larger scales of the
order of few 100~kpc up to Mpc, there are a vast number of
observations of radio emission coming from radio mini halos in the
center of cooling flow clusters, radio relics in the periphery of
clusters \cite{2004rcfg.proc..335K} as well smooth giant radio halos
that have been observed in more than 50 clusters
\cite{2003ASPC..301..143F,2008SSRv..134...93F}. Most of this observed
radio emission is expected to emerge from synchrotron emitting
cosmic-ray electrons whose precise origin in especially relics and
giant radio halos is still not settled. However, due to the smoothness
of the extended emission and the observed power-law slope it seems
quite natural that they are of hadronic origin. Especially since the
long cooling time of cosmic ray protons (CRs) allows for both an
extended and large population of CRs to build up over the cluster
history \cite{1997ApJ...487..529B}.

@@

\subsubsection{Inverse Compton}
\label{sect:IC}
The standard IC source function is given by
\cite{1979rpa..book.....R}:
\begin{eqnarray}
  q_{\rmn IC}(\eg, r) &=&  \frac{\dd^3 N_\gamma}{\dd V\,\dd t\,\dd \eg} = 
 \frac{3}{4}c\,\sigma_\rmn{T}
\int\dd \eph \frac{n_\rmn{ph}(\eph)}{\eph}\nonumber\\
&\times& \int \dd \ee\,\frac{\dd n_\e}{\dd \ee}(\ee,r)\,
 \frac{\left(m_\e c^2\right)^2}{\ee^2}G(\Gamma_\e,q)\,,\nonumber\\
  \label{eq:ICemiss}
\end{eqnarray}
where $\ee$ is the energy of the upscattering electrons and $\eph$ is
the energy of the background photon field. We represent the Thomson
cross-section with $\sigma_\rmn{T}$ and the full differential
Klein-Nishina (KN) cross-section is captured by
\cite{1970RvMP...42..237B}:
\begin{equation}
\label{eq:KN_spec}
G(\Gamma_\e,q) = 2q\ln{q}+(1+2q)(1-q)+ 
\frac{1}{2}\frac{\left(\Gamma_\e q\right)^2\left(1-q\right)}
     {1+\Gamma_\e q}\,,
\end{equation}
where
\begin{equation}
\Gamma_\e=\frac{4\eph \ee}{\left(m_\e c^2\right)^2}\,,\qquad \rmn{and} \qquad  
q=\frac{\eg}{\Gamma_\e\left(\ee-\eg\right)}\,.
\end{equation}
The full KN cross-section accounts for the less efficient energy
transfer between the photon and electron and results in a steepening
of the IC gamma-ray spectrum. In the low energy Thomson regime the IC
spectrum $F_\gamma\sim E_\gamma^{-(\alpha_\e-1)/2}$, however when
$\Gamma_\e \gg 1$ the IC spectrum steepens due to the KN suppression
to $\eg^{-\alpha_\e}\log(\eg)$. Here we denote the steady
state electon spectrum by $\alpha_\e$.

@@

\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/SB.photon.eps}
\caption{\it Spatial dependence of light from stars and dust. We show
  2D surface brightness profiles in the $r$-band ($\sim 1\,\ev$)
  obtained from stacked clusters in the Sloan Digital Sky Survey
  (SDSS) at the redshift $z \sim 0.25$ \cite{2005MNRAS.358..949Z}. The
  brightness profiles have been weighted with the $\rvir$ normalized
  area inside radius $r$, and trace the radial dependence of the
  luminosity from stars and galaxies. The crosses show the total
  observed starlight including the diffuse intracluster (ICL), the
  galaxies, and the brightest cluster galaxy (BCG) in the center of
  the cluster. The diamonds denote the observed starlight from the
  ICL and the BCG. The solid line show the fit to the data of the
  total light including the ICL, the BCG, and the galaxies, while the
  smooth component is represented by the dashed line that is fitted to
  only the ICL and the BCG.}
 \label{fig:SD_spatial}
\end{figure}

@@

The normalization constant for the starlight $N_\stars(\lx)$ is derived
in the following three steps: (1) We use the average apparent magnitude
in the $(r+i)$-band $m_{(r+i),0.25}\approx 15.5$
\cite{2005MNRAS.358..949Z} of a cluster with the average mass
$\mvir=4.0\times10^{13}\msun$ at a redshift $z=0.25$ to derive the
luminosity from the stars:
\begin{equation}
L_\stars=10^{\frac{m_\odot-m_{(r+i),0.25}}{2.5}}
\left(\frac{d_\clu}{d_\odot}\right)^2 L_\odot
\approx 7.1\times10^{44}\,\rmn{erg}\,\rmn{s}^{-1}\,,
\label{eq:L_star}
\end{equation}
where the apparent magnitude of the sun $m_\odot\approx -27.1$, the
distance to the sun $d_\odot=4.85\times10^{-6}\,\rmn{pc}$, the
luminosity of the sun $L_\odot=3.85\times10^{34}\,\rmn{erg\,s}^{-1}$,
and the distance to the cluster $d_\clu=1.26\times10^9\,\rmn{pc}$. (2)
We use the IR to X-ray scaling relation in Eq.~(\ref{eq:SD-X}) to find
a more general form of the luminosity from stars in
Eq.~(\ref{eq:L_star}) that is valid for an arbitrary cluster mass. The
cluster X-ray luminosities are found in the extended HIFLUGCS
catalogue \cite{2002ApJ...567..716R} to which we have added the Virgo
cluster. (3) The normalization constant for the stars is
determined from the above steps together with Eq.~(\ref{eq:E_SD})
where we integrate over the cluster volume and spectral distribution
of the stars. It is given by
\begin{equation}
 N_\stars(\lx) = 1.2\times10^{-9}\,
\left(\frac{\lx}{10^{44}\rmn{erg\,s}^{-1}}\right)^{0.53}\,.
\label{eq:N_stars}
\end{equation}
Note that since the luminosities include the contribution from
galaxies, we have integrated the spatial distribution including
galaxies to derive the normalization. Once the normalization is fixed,
the SD energy densities are derived from the spatial distribution
where galaxies have been excluded.

The normalization constant for the dust $N_\dust(\lx)$ is derived in the
following four steps: (1) We use a cluster with the the average mass
$\mvir=4.0\times10^{13}\msun$ and estimate the X-ray luminosity from
the X-ray to mass scaling relations derived for the Representative
XMM-Newton Cluster Structure Survey (REXCESS) in
\cite{2009A&A...498..361P}:
\begin{equation}
\lx=1.18\times 10^{44}\,\rmn{erg\,s}^{-1}
\left(\frac{M_{200}}{10^{14}\msun}\right)^{1.81}\,,
\end{equation}
where we have used that $M_\rmn{200} \approx 1.58\,M_\rmn{500}$ for a
small cluster \footnote{We derive the factor 1.58 using
  $M_\rmn{200}/M_\rmn{500}$ for the clusters in the HIFLUGCS catalogue
  with a mass similar to $\mvir=4.0\times10^{13}\msun$ that we use to
  derive the luminosity from dust.}. (2) From the IR to X-ray scaling
relation in Eq.~(\ref{eq:SD-X}) we can now estimate calculate the
luminosity from the dust for a $\mvir=4.0\times10^{13}\msun$ cluster;
$L_\dust=4.7\times10^{44}\,\rmn{erg\,s}^{-1}$. (3) We generalize
Eq.~(\ref{eq:L_star}) to an arbitrary cluster mass using
Eq.~(\ref{eq:SD-X}) together with the cluster X-ray luminosities in
the extended HIFLUGCS catalogue. (4) The normalization constant for
the dust can now be determined from from the above steps together with
Eq.~(\ref{eq:E_SD}) where we integrate over the cluster volume and
spectral distribution of the dust. It is given by
\begin{equation}
 N_\dust(\lx) = 1.4\times10^{-9}\,
\left(\frac{\lx}{10^{44}\rmn{erg\,s}^{-1}}\right)^{0.53}\,.
\label{eq:N_stars}
\end{equation}
Note that since the luminosities include the contribution from
galaxies, we have integrated the spatial distribution including
galaxies to derive the normalization. Once the normalization is fixed,
the SD energy densities are derived from the spatial distribution
where galaxies have been excluded.

@@

\section{Spectral gamma-ray distribution}
Spectrally resolved indirect DM searches have the advantage of probing
different DM models through their characteristic spectral
distributions. In this section we focus on the gamma-ray flux from
clusters its spectral distribution. We use both a leptophilic model
that give rise to substantial amounts of electrons and positrons that
IC upscatter background radiation fields to high energies as well as a
few supersymmetric DM models with a high gamma-ray yield in the form of
continuum emission. In addition to the annihilating DM, we estimate
the gamma-ray flux induced by shock accelerated CR protons.

In Fig.~\ref{fig:diff_BM} we show the differential flux from the
Fornax cluster which is one of the best clusters for inderect DM
serches due to its relative high DM fluxes. We derive the flux from
four different supersymmetric BM models and compare the flux to both
the emission induced by CRs as well as to the differential gamma-ray
upper limits set by Fermi-LAT after 18 months of observations
\cite{2010ApJ...717L..71A}. The upper limits are not violated,
although the predicted DM flux that is dominated by the continuum
emission from $\Kp$ and $\Ip$ (shown in the left panels) are close to
the upper limits and can potentially be probed by the Fermi 3 year
data. Furthermore, the expected masking gamma-ray signal induced by
CRs is about a factor 10 below the DM continuum flux from the $\Kp$
and $\Ip$ BM models aound 10~GeV. It is also important to note that IC
emission from upscattered CMB and SD photons is at least a factor 10
lower than the expected flux from CRs above 100~MeV, making it very
hard to distinguish such a signal from the foreground due to the
similar spectral index. Below the pion bump, however, we expect the IC
emission coming from shock accelerated electrons to be dominating
\cite{2010MNRAS.409..449P}. Hence the IC emission from SUSY DM in
clusters can for most applications safeley be neglected compared to
the CRs and continuum emission.

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
 \includegraphics[width=0.49\columnwidth]{figures/flux.BMcompI.v9.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompJ.v9.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompK.v9.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompJs.v9.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\caption{\it Comparing the differential flux from different models: we
  show the continuum emission from DM benchmark (BM) models (blue
  dashed), electrons and positrons from DM BM models that inverse
  Compton upscattere both CMB photons (orange dash-dotted) as well as
  dust and star photons (green dotted), and CR induced gamma-ray
  emission (red solid). Each panel is associated with an individual DM
  BM model; upper left $\Ip$, lower left $\Kp$, upper right $\Jp$, and
  lower right $\Js$. The emission is calculated for the Fornax cluster
  using a point spread function of $0.1\deg$ and a boost from
  substructures of ...  We find bright prospects for detecting the
  continuum emission that is dominating the total emission in the GeV
  energy range. Also note that, above 100~MeV the total inverse
  Compton emission is at least a factor 10 smaller than the emission
  expected from both the continuum emission and the emission coming
  from CRs.}
 \label{fig:diff_BM}
\end{minipage}
\end{figure*}

In Fig.~\ref{fig:IR_comp} we show the total IC emission as well as its
individual contributions from different IC upscattered radiation
fields. The left panel show the gamma-ray emission from the LP model
and the right panel from the $\Kp$ BM model, for comparison we
overplot the emission expected from the CRs. We find that the IC
upscattered CMB pfortons for the LP model overproduce the upper
gamma-ray limits set by Fermi. This puts tight constraints on the
boost from substructures as well as the Sommerfeld enhancement.

.........
Fermi can put tight limits on IC-CMB, much harder for IACTs to
constrain the flux due to the relative small SD flux and the lack of
substructures in the central cluster region
